### 2016-08-31 R/ITHIM.Rmd
Working on visulization of mean matrices
There's a bug here right now that needs to be fixed.
Visualiztion of Mean Matrices, Trip Speeds and Regional vs. National DALYs

```{r parsely}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

@@

```{r loadData}
NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
NHTS.trips <- readRDS(file = "./data/NHTS.trips.rds")
```

@@

```{r summary, fig.width = 15}
summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "walk"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "cycle"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "other"), TRVLCMIN))

walkHours <- with(subset(NHTS.trips, {
    AT == "walk" #& TRVLCMIN > 5 & TRPMILES > 2/9
}), TRVLCMIN/60)

walkMiles <- with(subset(NHTS.trips, {
    AT == "walk" #& TRVLCMIN > 5 & TRPMILES > 2/9
}), TRPMILES)

cycleHours <- with(subset(NHTS.trips, {
      AT == "cycle" #& TRVLCMIN > 5 & TRPMILES > 2/9
}), TRVLCMIN/60)

cycleMiles <- with(subset(NHTS.trips, {
      AT == "cycle" #& TRVLCMIN > 5 & TRPMILES > 2/9
}), TRPMILES)

@@

Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, mode = "walk")
Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, mode = "cycle")
D <- rbind(Dwalk,Dcycle)

@@

g <- ggplot(D, aes(Hours, Miles, colour = mode)) + geom_point()
g <- g + geom_abline(aes(slope = speed, intercept = intercept, colour = mode), Dspeed) + facet_grid(. ~ mode)
g + scale_y_log10() + scale_x_log10() + theme_bw()
```

@@

```{r subset}
NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
```

@@

```{r split}
listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
```

@@

```{r count}
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 0.01, x) )
RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 0.01, x) )
RwtList <- lapply( RwtList, function(x) x/x[3,2] )
RctList <- lapply( RctList, function(x) x/x[3,2] )
muwtList <- lapply( meanWalkList, function(x) x[3,2] )
muctList <- lapply( meanCycleList, function(x) x[3,2] )
```

@@

```{r ITHIMObjs, eval = FALSE}

ITHIMList <- list()

ITHIM.default <- createITHIM(vision = "baseline", region = "national")

for (HHS in paste0("HHS",1:10)){
  ITHIMList[[HHS]] <- ITHIM.default
  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muwt", muwtList[[HHS]])
  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muct", muctList[[HHS]])
  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rwt", RwtList[[HHS]])
  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rct", RctList[[HHS]])
  ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
}

@@

```{r CRA, eval = FALSE}

>>>>

```{r parsely, eval = FALSE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

@@

```{r loadData, eval = FALSE}
NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
NHTS.trips <- readRDS(file = "./data/NHTS.trips.rds")
```

@@

```{r summary, fig.width = 20, fig.height = 10, eval = TRUE}
summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "walk"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "cycle"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "other"), TRVLCMIN))

walkHours <- with(subset(NHTS.trips, {
    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVLCMIN*60 < 8 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 0.5
}), TRVLCMIN/60)

walkMiles <- with(subset(NHTS.trips, {
    AT == "walk" & TRPMILES < 10 & TRPMILES/TRVLCMIN*60 < 8 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 0.5
}), TRPMILES)

cycleHours <- with(subset(NHTS.trips, {
      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVLCMIN*60 < 25 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 1
}), TRVLCMIN/60)

cycleMiles <- with(subset(NHTS.trips, {
      AT == "cycle" & TRPMILES < 25 & TRPMILES/TRVLCMIN*60 < 25 & TRVLCMIN > 5 #& TRPMILES/TRVLCMIN*60 > 1
}), TRPMILES)

@@

Dwalk <- data.frame(Hours = walkHours, Miles = walkMiles, Speed = walkMiles/walkHours, mode = "walk")
Dcycle <- data.frame(Hours = cycleHours, Miles = cycleMiles, Speed = cycleMiles/cycleHours, mode = "cycle")

D <- rbind(Dwalk,Dcycle)

@@

g <- ggplot(D, aes(Hours, Miles, colour = mode)) + geom_point()
g <- g + geom_abline(aes(slope = speed, intercept = intercept, colour = mode), Dspeed) + facet_grid(. ~ mode)
g + scale_y_log10() + scale_x_log10() + theme_bw()

g <- ggplot(D, aes(Speed, ..density.., fill = mode)) + geom_histogram(breaks = seq(0,25,1)) + theme_bw() + scale_x_continuous(name="Speed") + scale_y_continuous(name="Density") + facet_grid(. ~ mode)
g

@@

```{r subset, eval = TRUE, fig.width = 10}
NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
barplot(with(NHTS,table(HHS)))
```

@@

```{r split, eval = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
```

@@

```{r count, eval = FALSE}
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 0.01, x) )
RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 0.01, x) )
RwtList <- lapply( RwtList, function(x) x/x[3,2] )
RctList <- lapply( RctList, function(x) x/x[3,2] )
muwtList <- lapply( meanWalkList, function(x) x[3,2] )
muctList <- lapply( meanCycleList, function(x) x[3,2] )
```

@@

```{r foobar, eval = TRUE}
sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
    return(sum(unlist(compareModels(ITHIM.baseline, ITHIM.scenario)$daly.delta)))
    }
```

```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10}
compute <- FALSE

if(compute){

DALYsTotal <- data.frame()

for( muNonTravel in seq(1,75,length.out = 20) ){

  ITHIMList <- list()
  rm(DALYs)
  ITHIM.default <- createITHIM(vision = "baseline", region = "national")

  for (HHS in paste0("HHS",1:10)){
    ITHIMList[[HHS]] <- ITHIM.default
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muwt", muwtList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muct", muctList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rwt", RwtList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "Rct", RctList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "GBD", GBDList[[HHS]])
    ITHIMList[[HHS]] <- updateITHIM(ITHIMList[[HHS]], "muNonTravel", muNonTravel)
  }

  ITHIM.national <- createITHIM(vision = "baseline", region = "national")
  ITHIM.national <- updateITHIM(ITHIM.national, "muNonTravel", muNonTravel)
  ITHIM.national <- updateITHIM(ITHIM.national, "muwt", 49.63)
  ITHIM.national <- updateITHIM(ITHIM.national, "muct", 2.50)
  DALYs <- lapply(ITHIMList, sumDALY, ITHIM.baseline = ITHIM.national)
  DALYs <- data.frame(HHS = names(DALYs), muNonTravel = muNonTravel, DALY = unlist(DALYs))
  DALYsTotal <- rbind(DALYsTotal, DALYs)
}
}else{
DALYsTotal <- readRDS(file = "./data/DALYsTotal.rds")
}
p <- ggplot(DALYsTotal, aes(x = muNonTravel, y = DALY, colour = HHS))
p  + geom_line() + theme_bw() + scale_x_discrete(name="Non-Travel METs mean") + scale_y_continuous(name="Change in DALY (National)") + facet_grid(. ~ HHS)
```

### Visualization of trip time matrices

```{r meanMatrices, eval = TRUE, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE}

ITHIMList <- c(ITHIMList, list("Natl" = ITHIM.national))

meanMatrix <- function(ITHIM, mode = "walk"){
   if(mode == "walk"){
      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rwt*muwt)))
}else{
      return(with(ITHIM$parameters, data.frame(ageClass = factor(1:8), mode = mode, Rct*muct)))
}
}

meanMatrixListWalk <- lapply(ITHIMList, meanMatrix, mode = "walk")
meanMatrixListCycle <- lapply(ITHIMList, meanMatrix, mode = "cycle")

D <- melt(c(meanMatrixListWalk,meanMatrixListCycle))

names(D) <- c("ageClass", "mode", "sex", "time", "HHS")

#D <- within(D, HHS <- factor(HHS, levels = c(paste0("HHS",1:10),"Natl")))

p1 <- ggplot(D, aes(x = ageClass, y = time, fill = sex)) + geom_bar(stat= "identity", position = "dodge") + facet_grid( mode ~ HHS)

p1

@@

```{r CRA, eval = TRUE}
